Compositional Quantitative Resource Analysis of C Programs

### 2014-06-10 paper/main.tex
more examples in paper

In this appendix we provide a non-exhaustive catalog of classes of
programs that can be automatically analyzed by our system.  For
simplicity, we use cost metric that counts cost $1$ for comparisons
and assignments and cost $0$ for all other operations.  Of course, the
examples can also be analyzed with any other cost metric.

@@

@@

>>>>

In this appendix we provide a non-exhaustive catalog of classes of
programs that can be automatically analyzed by our system.  For
simplicity, we use cost metric that counts cost $1$ for assignments
and cost $0$ for all other operations.  Of course, the examples can
also be analyzed with any other cost metric.

@@

  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  while (i>100) {
    i--;
  }
  i=i+k+50;
  while (i>=0) {
    i--;
  }
   \end{lstlisting}

$52 + |[-1,i]| + |[0,k]|$
\\[.7\baselineskip]
      {\bf t19}
    \end{center}
  \end{minipage}%
%
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  while (x<y) {
    x=x+1;
  }
  while (y<x) {
    y=y+1;
  }
   \end{lstlisting}

$|[x,y]|+|[y,x]|$
\\[.7\baselineskip]
      {\bf t20}
    \end{center}
  \end{minipage}%
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  while (x>0) {
    x=x-1;
    t=x;
    x=y;
    y=t;
  }
   \end{lstlisting}

$4|[0,x]|+4|[0,y]|$
\\[.7\baselineskip]
      {\bf t30}
    \end{center}
  \end{minipage}

   \caption{Amortization and Compositionality (c)}
  \label{fig:cat1c}
\end{figure*}

@@

Example \emph{t19} demonstrates another nice property that demonstates
the compositionaliy of the analysis.  The program consists of two
loops that decrement a variable $i$.  In the first loop, $i$ is
decremented down to 100 and in the second loop $i$ is increment
further down to $-1$.  However, between the loops we assign
$i=i+k+50$.  So in total the program performs $52 + |[-1,i]| +
|[0,k]|$ increments.  Our analysis finds this tight bound because our
amortized analysis naturally takes into account the relation between
the two loops.  Other techniques might derive a more conservative
bound such as $52 + |[-1,i]| + |[0,k]| + |[100,i]|$.

Example \emph{t20} shows how we can handle programs in which bounds
contain absolute values like $|x-y|$.  The first loop body is only
executed if $x<y$ and the second loop body is only executed if $y<x$.
The analyzers finds a tight bound.

At first sight, example \emph{t30} appears to be a simple loop that
decrements the variable $x$ down to zero.  However, a closer look
reveals that the loop actually decrements both input variables $x$ and
$y$ down to zero before terminating.  In the loop body, first $x$ is
decremented by one.  Then the values of the variables $x$ and $y$ are
switched using the local variable $t$ as a buffer.  Out analysis
infers the tight bound $4|[0,x]|+4|[0,y]|$.

\begin{figure*}[t!]
 \setlength{\progwidth}{.24\linewidth}
  \centering
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  while (n>x) {
    if (m>y) 
      y = y+1;
    else
      x = x+1;
  }
   \end{lstlisting}

$|[x, n]| + |[y, m]|$
\\[.7\baselineskip]
      {\bf t11}
    \end{center}
  \end{minipage}
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
 XXX
   \end{lstlisting}

$XXX$
\\[.7\baselineskip]
      {\bf t12}
    \end{center}
  \end{minipage}
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
 XXX
   \end{lstlisting}

$XXX$
\\[.7\baselineskip]
      {\bf t22}
    \end{center}
  \end{minipage}

   \caption{Examples from \cite{GulwaniMC09}}
  \label{fig:cat2a}
\end{figure*}

   \caption{Examples from [GulwaniPLDI09] and [GulwaniPLDI10]}
  \label{fig:cat2b}
\end{figure*}

\paragraph{From  the Literature}

Example \emph{t11} has been given in~\cite{GulwaniMC09}.  In the loop,
we increment

\begin{figure*}[t!]
 \setlength{\progwidth}{.24\linewidth}
  \centering
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
 XXX
   \end{lstlisting}

$XXX$
\\[.7\baselineskip]
      {\bf Knuth-Morris-Pratt}
    \end{center}
  \end{minipage}
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
 XXX
   \end{lstlisting}

$XXX$
\\[.7\baselineskip]
      {\bf GCD}
    \end{center}
  \end{minipage}

   \caption{Well-Known Algorithms}
  \label{fig:cat3}
\end{figure*}

\paragraph{Well-Known Algorithms}
